Polynomial phase estimation by phase unwrapping

### 2013-02-23 paper.tex
more paper edits

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping} (LSU) estimator~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam2010thesis}.  It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  
A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999,Agrell2002} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$.  Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered has significant structure, and may admit fast nearest point algorithms.  This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.  Although polynomial time, the algorithms are still computationally demanding in practice.  Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.  In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} for approximating the LSU estimator.  We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large (approximately $N$ less than $1000$).  %The $K$-best method is motivated by sequential decoding algorithms and, in the context of polynomial phase estimation, bares some resemblance to the Bayesian unwrapping algorithm studied by Morelande~\cite{Morelande_bayes_unwrapping_2009_tsp}.
%We show by Montecarlo simulation that the LSU estimator approximated using the $K$-best method is statistically superior to state-of-the-art estimators from the literature including the HAF and CPF.  We also compare the computational performance of the estimators and find that BLERG.

@@

An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see inparticular the proof of Lemma~1 and the remark that follows).  To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.  This has apparently lead to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.  This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occuring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.  The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.  The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.  For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to approximately 3 times the CRB as the SNR approaches zero.  Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.  It is worth noting that this factor of 3 times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.

@@

A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}. The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.  On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.  Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small \cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}. One such algorithm is the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999}.  With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.  Algorithms that approximate a nearest point have also been studied.  One example is Babai's nearest plane algorithm \cite{Babai1986}.  Yet another approximate algorithm is the $K$-best algorithm studied in~\cite{Zhan2006_K_best_sphere_decoder} that operates similarly to the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} used in coding theory.  The $K$-best algorithm is more expensive that Babai's algorithm, but typically produces a better approximation.  In this paper we make use of the $K$-best algorithm for approximating the LSU estimator.  We will find that it produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).   We will not detail the workings of the $K$-best algorithm in this paper.  A thorough description is given in~\cite{Zhan2006_K_best_sphere_decoder}.

@@

This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.  The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.  In all our simulations the unknown amplitude is $\rho = 1$ and $N = 199$.  The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.  The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.  The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.  The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.  In all simulations the LSU estimator is approximated using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder}.

Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.  All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.  The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.  Provided the noise variance is small enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.  Using Theorem~\ref{thm:asymp_proof} we predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and approximately $3$ as SNR approaches zero~\cite{Quinn2009_dasp_phase_only_information_loss}.  Contrastingly, the ratio between the CRB and asymptotic variance of the HAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.  %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.

@@

Figures~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.  In this case the LSU estimator reaches a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.  In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF and CPF-HAF estimators.  As in the case where $m=3$, the region of coefficients suitable for the HAF and CPF-HAF estimators is much smaller that the identifiable region $B$.  The effect becomes more servere as $m$ increases.  When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.  This constitutes is a dramatic reduction in the range of detectable polynomial phase signals. Simulations have been run with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$, and in this case we obtain results resembling those in Figure~\ref{plot:m3big}.  The performance of the LSU estimator remains unchanged, but the HAF and CPF-HAF estimators fail completely.

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-1.mps} 
  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.} 
  		\label{plot:m3small} 
 \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-2.mps} 
  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
  		\label{plot:m3big} 
 \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-3.mps} 
  		\caption{Polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
  		\label{plot:m3bigzw} 
 \end{figure}

\begin{figure}[t] 
   	\centering 
  		\includegraphics{code/data/plot5-1.mps} 
  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}   \right]$.} 
  		\label{plot:m5} 
\end{figure}

\begin{figure}[t] 
   	\centering 
  		\includegraphics{code/data/plot5-2.mps} 
  		\caption{Polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.} 
  		\label{plot:m5} 
\end{figure}

@@

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). It has been shown that the LSU estimator is strongly consistent and asymptotically normally distributed. %Simulations are used to show that the LSU has near maximum likelihood performance in the case that the noise is additive white and Gaussian.  
The LSU estimator can be computed by finding a nearest lattice point in a lattice~\cite{McKilliam2009asilomar_polyest_lattice}.  %Here, we consider using the $K$-best method~\cite{Zhan2006_K_best_sphere_decoder} to achieve this in reasonable time when the number of observations, $N$, is less than about 300.  
Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.  Instead, we considered some general purpose algorithms, the \term{sphere decoder}, \term{Babai's nearest plane algorithm} and the \term{$K$-best algorithm}. The sphere decoder and $K$-best algorithm result in accurate estimators.

The major outstanding question is whether faster nearest point algorithms exist for these specific lattices.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.

>>>>

In this paper we consider the estimator that results from unwrapping the phase in a least squares manner.  We call this the \emph{least squares unwrapping} (LSU) estimator~\cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam2010thesis}.  It was shown in \cite{McKilliam2009asilomar_polyest_lattice, McKilliamFrequencyEstimationByPhaseUnwrapping2009} that the LSU estimator can be represented as a \emph{nearest lattice point problem}~\cite{Agrell2002}, and Monte-Carlo simulations were used to show the LSU estimator's favourable statistical performance. %Moreover, it was noticed that the LSU appears to work correctly for coefficients anywhere in the identifiable region.  
A drawback of the LSU estimator is that computing a nearest lattice point is, in general, computationally difficult.  In \cite{McKilliam2009asilomar_polyest_lattice}, two standard techniques were considered, the \emph{sphere decoder}~\cite{Pohst_sphere_decoder_1981,Viterbo_sphere_decoder_1999,Agrell2002} that exactly computes a nearest lattice point, and \emph{Babai's nearest plane algorithm}~\cite{Babai1986} that only produces an approximation. The sphere decoder was observed to have excellent statistical performance but can only be computed efficiently for small $N$ less than about $60$.  Babai's algorithm is computationally simple, but its statistical performance was observed to be comparatively poor at low signal to noise ratio (SNR). A major point of interest is that the lattices considered has significant structure, and may admit fast nearest point algorithms.  This has been studied in~\cite{McKilliam2010thesis} where polynomial time algorithms were found that compute the nearest point exactly.  Although polynomial time, the algorithms are still computationally demanding in practice.  Nevertheless fast (exact or approximate) algorithms may exist for these special lattices, that are yet to be discovered.  In this paper we apply another general purpose approximate nearest point algorithm, called the $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} for approximating the LSU estimator.  We show that this provides near sphere decoder (exact) performance, but can be computed in a reasonable amount of time if $N$ is not too large (approximately $N$ less than $1000$).  %The $K$-best method is motivated by sequential decoding algorithms and, in the context of polynomial phase estimation, bares some resemblance to the Bayesian unwrapping algorithm studied by Morelande~\cite{Morelande_bayes_unwrapping_2009_tsp}.
%We show by Montecarlo simulation that the LSU estimator approximated using the $K$-best method is statistically superior to state-of-the-art estimators from the literature including the HAF and CPF.  We also compare the computational performance of the estimators and find that BLERG.

@@

An analysis of the HAF estimator under the assumption that the noise $\{X_n\}$ is white and Gaussian and without the `small noise' assumption has been made in~\cite{Porat_asympt_HAF_DPT_1996} (see inparticular the proof of Lemma~1 and the remark that follows).  To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.  This has apparently lead to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.  This misconception would appear to be motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occuring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.  The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.  The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss.  For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that the asymptotic variance of the LSU estimator converges to the Cramer-Rao lower bound (CRB) as the SNR becomes large, and converges to $\tfrac{2}{3}\pi$ times the CRB as the SNR approaches zero.  Related results have been presented previously~\cite{Quinn2009_dasp_phase_only_information_loss}.  It is worth noting that this factor of $\tfrac{2}{3}\pi$ times the CRB at low SNR is better than popular estimators such as the HAF and CPF, for which the ratio between the asymptotic variance and the CRB diverges as the SNR approaches zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.

@@

A fundamental problem in lattice theory is the \emph{nearest lattice point problem}~\cite{Agrell2002}. The nearest lattice point problem is: Given $\ybf\in\reals^n$ and some lattice $\Lambda \subset \reals^n$, find a lattice point $\xbf \in \Lambda$ such that the Euclidean distance between $\ybf$ and $\xbf$ is minimised, that is, solve $\arg\min_{\xbf \in \Lambda} \|\ybf - \xbf\|^2.$  The nearest lattice point problem is known to be NP-hard under certain conditions when the lattice itself is considered as an additional input parameter \cite{micciancio_hardness_2001, AjtaiShortestVecProbNPHard1998}.  On the other hand,  fast algorithms that compute a nearest point often exist for specific lattices with known structure~\cite{SPLAG,Conway1982FastQuantDec,McKilliam2009CoxeterLattices,McKilliam2008}.  Even in the general case, where the lattice has no known structure, algorithms exist that can compute a nearest lattice point in reasonable time if the dimension is small~\cite{Agrell2002, Viterbo_sphere_decoder_1999, Pohst_sphere_decoder_1981}. These algorithms typically operate using a tree search to enumerate all lattice points inside a ball of radius sufficiently large to contain a nearest point.  The algorithms vary in the order that branches in the tree are searched.  A popular approach due to Schnorr and Eucher~\cite{schnorr_euchner_sd_1994,Agrell2002} searches branches in a greedy order, and this is the approach we implement here.  We refer to this algorithm as the~\emph{sphere decoder} in this paper, although other variants of the algorithm also go by this name.

With modern computers, the sphere decoder is computationally viable if the dimension of the lattice is less than approximately $60$.  Algorithms that approximate a nearest point have also been studied.  One example is Babai's nearest plane algorithm \cite{Babai1986}.  Another pragmatic approximate approach is to employ the sphere decoder, but prune branches in the search tree to ensure that there are atmost a finite number, say $K$, remaining branches at any time.  This approach has been studied previously~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009} and resembles what is called the sequential $M$-algorithm~\citep{Anderson1984_seq_coding_alg} from coding theory.  The approaches vary in how branches are pruned.  In our implementation we prune branches using the same greedy metric as that used by the Schnorr and Euchner sphere deocder~\cite{schnorr_euchner_sd_1994,Agrell2002} for deciding the order branches are searched.  Following~\cite{Zhan2006_K_best_sphere_decoder} we refer to this as \emph{$K$-best algorithm}.  We will find that this algorithm produces excellent estimates of polynomial phase coefficients and can be computed in a reasonable amount of time for moderate numbers of observations ($N$ less that about 1000).

@@

This section describes the results of Monte-Carlo simulations with the least squares unwrapping (LSU) estimator, the high order ambiguity function (HAF) estimator~\cite{Peleg_DPT_1995}, the cubic phase function (CPF) estimator~\cite{Oshea_cpf_2004} and the estimator that results from combining the HAF and CPF~\cite{Djurovic_haf_cpf_2012}.  The CPF estimator only applies to polynomial phase signals of order $m=3$ and when the number of observations $N$ is odd.  In all our simulations the unknown amplitude is $\rho = 1$ and $N = 199$.  The $X_1, \dots, X_N$ are pseudorandomly generated independent and identically distributed circularly symmetric complex Gaussian with variance $\expect\abs{X_1}^2 = \sigma_c^2$.  The corresponding signal to noise ratio (SNR) is $\rho^2 \sigma_c^{-2} = \sigma_c^{-2}$.  The number of replications of each Monte-Carlo experiment is $T = 2000$ to obtain estimates $\widehat{\mubf}_1, \dots, \widehat{\mubf}_T$ and the corresponding dealiased errors $\widehat{\lambdabf}_t = \dealias(\widehat{\mubf}_T - \tilde{\mubf})$ are computed.  The sample mean square error (MSE) of the $k$th coefficient is computed according to $\tfrac{1}{T}\sum_{t=1}^T \widehat{\lambda}_{k,t}^2$ where $\widehat{\lambda}_{k,t}$ is the $k$th element of $\widehat{\lambdabf}_t$.  In all simulations the LSU estimator is approximated using the $K$-best method described in Section~\ref{sec:lattice-theory}.  The maximum number of branches allowed is set to $K=20*N$.

Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.  All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.  The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.  Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.  Using Theorem~\ref{thm:asymp_proof} we can predict the ratio between the CRB and the asymptotic variance of the LSU estimator to approach 1 as SNR approaches $\infty$ and $\tfrac{2}{3}\pi$ as SNR approaches zero~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).  Contrastingly, the ratio between the CRB and asymptotic variance of the HAF and CPF estimators is known to diverge as SNR goes to zero~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.  %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.

@@

Figures~\ref{plot:m5} displays the sample MSE of the highest order coefficient $\mu_5$  for a polynomial phase signal of order $m=5$.  In this case the LSU estimator reaches a threshold at \unit[12]{dB}, the HAF estimator at \unit[18]{dB}, and the estimator that combines the CPF and HAF at \unit[10]{dB}.  In this figure the true coefficients are $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}  \right]$ and these are chosen specifically to satisfy the requirement of the HAF and CPF-HAF estimators.  As in the case where $m=3$, the region of coefficients suitable for the HAF and CPF-HAF estimators is much smaller that the identifiable region $B$.  The effect becomes more servere as $m$ increases.  When $m=5$ and $N=199$ the space of coefficients searched by the HAF and CPF-HAF estimators is approximately $10^{16}$ times smaller than the identifiable region.  This constitutes a dramatic reduction in the range of detectable polynomial phase signals. Figure~\ref{plot:m3big} displays results under the same conditions but with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}  \right]$.  The conclusion is similar to that from Figure~\ref{plot:m3big}.  The performance of the LSU estimator remains unchanged, but the HAF and CPF-HAF estimators fail completely.

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-1.mps} 
  		\caption{Sample MSE for highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}  \right]$.} 
  		\label{plot:m3small} 
 \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-2.mps} 
  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
  		\label{plot:m3big} 
 \end{figure}

\begin{figure}[p] 
   	\centering 
  		\includegraphics{code/data/plot3-3.mps} 
  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_3$ for a polynomial phase signal of order 3 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}  \right]$.} 
  		\label{plot:m3bigzw} 
 \end{figure}

\begin{figure}[t] 
   	\centering 
  		\includegraphics{code/data/plot5-1.mps} 
  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8N}, \tfrac{1}{24N^2}, \tfrac{1}{96N^3}, \tfrac{1}{480N^4}   \right]$.} 
  		\label{plot:m5} 
\end{figure}

\begin{figure}[t] 
   	\centering 
  		\includegraphics{code/data/plot5-2.mps} 
  		\caption{Sample MSE for the highest order coefficientl $\tilde{\mu}_5$ for a polynomial phase signal of order 5 with $\tilde{\mubf} = \left[ \tfrac{1}{4}, \tfrac{1}{4}, \tfrac{1}{8}, \tfrac{1}{24}, \tfrac{1}{96}, \tfrac{1}{480}   \right]$.} 
  		\label{plot:m5} 
\end{figure}

@@

This paper has considered the estimation of the coefficients of a noisy polynomial phase signal by least squares phase unwrapping (LSU). The LSU estimator was shown to be strongly consistent and asymptotically normally distributed under mild conditions on the distribution of the noise.  The estimator can be computed by finding a nearest lattice point in a lattice. Polynomial time nearest point algorithms for these lattices exist~\cite[Sec 4.3]{McKilliam2010thesis}, but these algorithms are not fast in practice.  We instead make used on the appromxiate $K$-best algorithm~\cite{Baro_list_sphere_dec_2003,Zhan2006_K_best_sphere_decoder,Mondal_Kbest_again_2009}.  This algorithm results in an accurate estimator and can be computed in a reasoable amount of time for moderately large numbers of obervations $N$.

A key property of the LSU estimator is that it functions correctly for polynomial phase parameters contained anywhere inside the identifiable region described in Section~\ref{sec:ident_aliasing}.  This is the maximum possible range of parameters.  By constrast popular estimators such as the HAF and CPF operate correctly only for parameters in a range much smaller than the identifiable region.  For this reason, the LSU estimator may be the only presently existing estimator that is viable for certain applications, such as interference supression~\cite{Djukanovic:2011:PMN:1950994.1951252,Djurovic_aliasing_2012}.

The major outstanding question is whether faster nearest point algorithms exist for these specific lattices, particularly for the case when $N$ is large.  Considering the excellent statistical performance (both theoretically and practically) of the LSU estimator, even fast \emph{approximate} nearest point algorithms are likely to prove useful for the estimation of polynomial phase signals.
